Legion: Expressing Locality and Independence with Logical Regions

### 2012-05-05 doc/sc2012/circuit.tex

The key decisions in a Legion program are how data is grouped into
regions and how  regions are {\em partitioned} into {\em
subregions}.  
%
%FIXME
% the following sentence oversimplifies; it's one of the goals, and regions tell us
% useful things about dependent computations as well
%
The goal is to pick an organization that makes explicit
which computations are independent.  
A {\tt Circuit}
has two regions: collections of nodes and wires (line
3 of Listing~\ref{lst:code_ex}).\footnote{Note that pointers to region values declare the region to which they point.  For
example, the definition of {\tt Wire} (line 2) is parametrized on the region
{\tt rn} to which the {\tt Node} pointers in fields {\tt in\_nodes}
and {\tt out\_nodes} point.}
An efficient parallel
implementation breaks this unstructured graph into pieces that can be
processed (mostly) independently. An appropriate
region organization makes explicit which nodes and wires are
involved in intra-piece computation and, where wires connect different pieces,
which are involved in inter-piece computation.

% FIXME   We add here a description of how the colors print when printed greyscale.
%
Figure~\ref{sfig:part_fig:pvs} shows how the nodes
in a small graph might be split into three pieces.  Blue nodes,
attached by wires only to nodes in the same piece, are 
{\em private} to the piece.  Orange nodes, on the boundary of a
piece, are {\em shared} with (connected to) other pieces.
In the simulation, computations on the
private nodes of different pieces are independent, while
computations on the shared nodes require communication.  To make
this explicit in the program, we partition the nodes region into
private and shared subregions (line 18).  To partition a region, we
provide a {\em coloring}, which is a relation between the elements of
a region and a set of colors.  A partition is an object which given a
coloring and a region $r$ contains for each color $c$ a subregion of
$r$ containing all the elements of $r$ colored $c$. Note that the
partition into shared and private nodes is disjoint because each node has one color.

@@

%
% FIXME
% Sean wants to use the term task through this paragraph; however we haven't defined it yet.
% I think it is OK as is since the term is introduced at the end of the paragraph.
%
Line 9 declares the main simulator function, which specifies the regions it 
accesses and the privileges and coherence it requires of those regions.
The {\tt RWE} annotation specifies that
the regions {\tt c.r\_all\_nodes} and {\tt c.r\_all\_wires}
are accessed with read-write privileges and {\em exclusive} coherence (i.e., no other
task can access these two regions concurrently or be reordered around this
task if they use either region).  The simulation 
reads and writes all nodes and wires, and it must be done
with exclusive access to ensure correct results.  Privileges specify what
the function can do with the regions; coherence specifies what other
functions can do with the regions concurrently.  Functions that
declare their accessed regions, privileges, and coherence are called {\em tasks}
and are the unit of parallel execution in Legion.

Lines 31-57 form the bulk of the actual simulation, performing three
passes over the circuit in each iteration.  Each pass loops over an
array of pieces (constructed on lines 27-30 from the partitions),
spawning a task for each piece.  There are no explicit requests for
parallel execution ({\tt spawn} simply indicates a task call and does
not require parallel execution) nor is there synchronization 
between the passes.  Which tasks can be run in
parallel within a pass and the required inter-pass dependencies are
determined automatically by the Legion runtime based on the region
access annotations on the task declarations.  The tasks spawned on
lines 32-34 are {\em subtasks} of the main {\tt simulate\_circuit}
task. A subtask can only access regions (or subregions) that its parent task
could access; furthermore, the subtask can only have permissions on a
region compatible with the parent's permissions.  
%This
%requirement plays an important role in our task scheduling algorithm
%(see Section~\ref{sec:soop}).

@@

%The two basic data types in the circuit simulation are {\tt Node}s and
%{\tt Wire}s declared on lines 1 and 2 of Listing~\ref{lst:code_ex}.
%As is standard in region-based systems, allocating a {\tt Node}, {\tt
%  Wire}, or any value in the heap requires naming the region
%in which the value is allocated.  For example, {\tt new Node@r} returns
%a reference to a new value of {\tt Node} type in region {\tt r}.  
Listing~\ref{lst:code_ex} illustrates one way of constructing partitioned
data structures in Legion: populate a region with data (the example
makes use of whatever data has been allocated by the caller of {\tt simulate\_circuit})
and then partition it into subregions.  One can also first partition an empty region into subregions
and then allocate data in the subregions. We use both approaches in our
applications (see Section~\ref{sec:exp}).

>>>>

The key decisions in a Legion program are how data is grouped into
regions and how  regions are {\em partitioned} into {\em
subregions}.  
%
%FIXME
% the following sentence oversimplifies; it's one of the goals, and regions tell us
% useful things about dependent computations as well
%
The goal is to pick an organization that makes explicit
which computations are independent.  
A {\tt Circuit}
has two regions: a collection of nodes and a collection of wires (line
3 of Listing~\ref{lst:code_ex}).\footnote{Note that all pointers declare the region to which they point.  For
example, the definition of {\tt Wire} (line 2) is parametrized on the region
{\tt rn} to which the {\tt Node} pointers in fields {\tt in\_nodes}
and {\tt out\_nodes} point.}
An efficient parallel
implementation breaks this unstructured graph into pieces that can be
processed (mostly) independently. An appropriate
region organization makes explicit which nodes and wires are
involved in intra-piece computation and, where wires connect different pieces,
which are involved in inter-piece computation.

% FIXME   We add here a description of how the colors print when printed greyscale.
%
Figure~\ref{sfig:part_fig:pvs} shows how the nodes
in a small graph might be split into three pieces.  Blue nodes,
attached by wires only to nodes in the same piece, are 
{\em private} to the piece.  Orange nodes, on the boundary of a
piece, are {\em shared} with (connected to) other pieces.
In the simulation, computations on the
private nodes of different pieces are independent, while
computations on the shared nodes require communication.  To make
this explicit in the program, we partition the nodes region into
private and shared subregions (line 18).  To partition a region, we
provide a {\em coloring}, which is a relation between the elements of
a region and a set of colors.  
%A partition is an object which given a
%coloring and a region $r$ contains for each color $c$ a subregion of
%$r$ containing all the elements of $r$ colored $c$. 
For each color $c$ in the coloring, the partition contains a subregion $r$
of the region being partitioned, with $r$ consisting of the elements colored $c$.  Note that the
partition into shared and private nodes is disjoint because each node has one color.

@@

%
% FIXME
% Sean wants to use the term task through this paragraph; however we haven't defined it yet.
% I think it is OK as is since the term is introduced at the end of the paragraph.
%
Line 9 declares the main simulator function, which specifies the regions it 
accesses and the privileges and coherence it requires of those regions.
The {\tt RWE} annotation specifies that
the regions {\tt c.r\_all\_nodes} and {\tt c.r\_all\_wires}
are accessed with read-write privileges and {\em exclusive} coherence (i.e., no other
task can access these two regions concurrently or be reordered around this
task if they use either region).  
%The simulation 
%reads and writes all nodes and wires, and it must be done
%with exclusive access to ensure correct results.  
Privileges specify what
the function can do with the regions; coherence specifies what other
functions can do with the regions concurrently.  Functions that
declare their accessed regions, privileges, and coherence are called {\em tasks}
and are the unit of parallel execution in Legion.

Lines 31-57 form the bulk of the actual simulation, performing three
passes over the circuit in each iteration.  Each pass loops over an
array of pieces (constructed on lines 27-30 from the partitions),
spawning a task for each piece.  There are no explicit requests for
parallel execution ({\tt spawn} simply indicates a task call and does
not mandate parallel execution) nor is there synchronization 
between the passes.  Which tasks can be run in
parallel within a pass and the required inter-pass dependencies are
determined automatically by the Legion runtime based on the region
access annotations on the task declarations.  The tasks spawned on
lines 32-34 are {\em subtasks} of the main {\tt simulate\_circuit}
task. A subtask can only access regions (or subregions) that its parent task
could access; furthermore, the subtask can only have permissions on a
region compatible with the parent's permissions.  
%This
%requirement plays an important role in our task scheduling algorithm
%(see Section~\ref{sec:soop}).

@@

%The two basic data types in the circuit simulation are {\tt Node}s and
%{\tt Wire}s declared on lines 1 and 2 of Listing~\ref{lst:code_ex}.
%As is standard in region-based systems, allocating a {\tt Node}, {\tt
%  Wire}, or any value in the heap requires naming the region
%in which the value is allocated.  For example, {\tt new Node@r} returns
%a reference to a new value of {\tt Node} type in region {\tt r}.  
Listing~\ref{lst:code_ex} illustrates one way of constructing partitioned
data structures in Legion: populate a region with data (the example
makes use of whatever data has been allocated by the caller of {\tt simulate\_circuit})
and then partition it into subregions.  One can also first partition an empty region into subregions
and then allocate data in the subregions. 
%We use both approaches in our
%applications (see Section~\ref{sec:exp}).

### 2012-05-05 doc/sc2012/soop.tex

Once a task $t$ has been mapped it enters the execution stage of the SOOP.  When all of
the operations (other tasks and copies) on which $t$ depends have completed, $t$ is
launched on the processor.  When $t$ spawns subtasks, they are recursively scheduled as siblings by
the SOOP using $t$ as the parent task and $t$'s region arguments as the roots of the region forest.

>>>>

After a task $t$ has been mapped it enters the execution stage of the SOOP.  When all of
the operations (other tasks and copies) on which $t$ depends have completed, $t$ is
launched on the processor it was mapped to.  When $t$ spawns subtasks, 
they are registered by the SOOP on which $t$ is executing, using $t$ as the parent task 
and $t$'s region arguments as the roots of the region forest.  Each child of $t$ traverses
the SOOP pipeline on the same processor as $t$, possibly being mapped to a 
SOOP instance on a different processor to execute.
